Difference in 2019 Canadian Federal Election if Everyone had Voted

### 2020-12-17 clean data scripts/clean_census_data.Rmd
model and result to be fixed

```{r,warning=FALSE, message=FALSE}
# Keep variables that we're interested
reduced_dat <- 
  gss %>% 
  select(sex,
         education,
         worked_last_week,
         place_birth_canada,
         province,
         income_family,
         age
         )
```

```{r}
# rename some variables 
census <- reduced_dat %>% rename(
  edu = education,
  income = income_family,
  employed = worked_last_week,
  bornin_canada = place_birth_canada)
```

@@

```{r}
#Create a .csv file
write.csv(census, "census_cleaned.csv")
```

>>>>

```{r,warning=FALSE, message=FALSE}
# Keep variables that we're interested
reduced_dat <- 
  gss %>% 
  select(sex,
         education,
         worked_last_week,
         place_birth_canada,
         province,
         age
         )
```

```{r}
# rename some variables 
census <- reduced_dat %>% rename(
  edu = education,
  employed = worked_last_week,
  bornin_canada = place_birth_canada)
```

@@

```{r}
#Create cell counts:
cell_counts <- census %>% group_by(province, sex, as.factor(age), employed, bornin_canada, edu) %>% count() %>% rename(age = "as.factor(age)")
#Create a .csv file
write.csv(cell_counts, "census_cleaned.csv")
```

### 2020-12-17 output/paper.Rmd
model and result to be fixed

---
title: "Difference in 2019 Canadian Federal Election if Everyone had Vote"
author: "Qiyue Zhang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    fig_width: 7
    fig_height: 4
abstract: |
  In the last Canadian Federal Election, the Liberal Party narrowly won the most of the seats in the parliament and its biggest adversary the Conservative Party has gained more seats than the last election. In this paper, we aim to identify how the 2019 Canadian Federal Election would have been different if "everyone" had voted. We ran a multilevel logistic regression model using observations from the results of online survey data provided by the 2019 Canadian Election Study (CES). To provide more robust conclusion, we utilize the regression with post-stratification method using 2017 General Social Survey (GSS) data provided by Statistics Canada. Our research finds out that Conservative actually wins the popular vote. Hence, our finding shows the importance of converting people to your cause, getting them to turn out to vote, and how the two are linked.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setup library
library(ggplot2)
library(dplyr)
library(knitr)
library(readr)
library(tidyverse)
library(broom)
library(lme4)
library(readr)
#import cleaned data use for analysis
survey <- read_csv("survey_cleaned.csv")
post <- read_csv("census_cleaned.csv")
#remove NA by na.omit()
survey <- na.omit(survey)
post <- na.omit(post)
```

@@

# 3 Model

# 4 Results

>>>>

---
title: "Difference in 2019 Canadian Federal Election if Everyone had Voted"
author: "Qiyue Zhang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    fig_width: 7
    fig_height: 4
abstract: |
  In the last Canadian Federal Election, the Liberal Party narrowly won the most of the seats in the parliament and its biggest adversary the Conservative Party has gained more seats than the last election. In this paper, we aim to identify how the 2019 Canadian Federal Election would have been different if "everyone" had voted. We ran a multilevel logistic regression model using observations from the results of online survey data provided by the 2019 Canadian Election Study (CES). To provide more robust conclusion, we utilize the regression with post-stratification method using 2017 General Social Survey (GSS) data provided by Statistics Canada. Our research finds out that Conservative actually wins the popular vote. Hence, our finding shows the importance of converting people to your cause, getting them to turn out to vote, and how the two are linked.

@@

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setup library
library(ggplot2)
library(dplyr)
library(knitr)
library(readr)
library(tidyverse)
library(broom)
library(lme4)
library(readr)
#import cleaned data use for analysis
survey <- read_csv("survey_cleaned.csv")
post <- read_csv("census_cleaned.csv")
#remove NA by na.omit()
survey <- na.omit(survey)
post <- na.omit(post)
#remove income variable no longer used for survey data
survey <- survey[,-8]
```

@@

# 3 Model
## 3.1 Multilevel Logistic Regression Model
```{r, include=FALSE}
# Construct the vote intention variable from survey data so that it is binary indicates by 1 and 0
survey$vote_Conservative[survey$votechoice=="Conservative Party"]<-1
survey$vote_Conservative[survey$votechoice=="Liberal Party"]<-0
```

We are interested in determining the popular vote outcome for the Conservative Party for the 2019 Canadian Federal Election by using a multilevel logistic regression based on the factors of sex, education, employment, born in Canada, and age group for different intercept for each province.

Multilevel logistic regression model: 
$$\log \left(\frac{\hat{p}_{ij}}{1-\hat{p}_{ij}}\right)={\beta}_{0}+{\beta}_{1} x_{ij}^{sexMale}+{\beta}_{2} x_{ij}^{educationed}+{\beta}_{3} x_{ij}^{employed}+{\beta}_{4} x_{ij}^{bornCA}$$
$$+{\beta}_{5} x_{ij}^{age30-44}+{\beta}_{6} x_{ij}^{age45-59}+{\beta}_{7} x_{ij}^{age60-74}+{\beta}_{8} x_{ij}^{age75above}$$

where $i=1,...,711$(number of observations), $j=1,..., 10$(number of provinces),

$p$ represents the probability of voting for the Conservative Party,

$\frac{p}{1-p}$ represents the odds of voting for the Conservative Party,

${\beta}_{1}$ coefficient represents the change in log odds for male respondents,

${\beta}_{2}$ coefficient represents the change in log odds for respondent who has received some college education,

${\beta}_{3}$ coefficient represents the change in log odds for respondent who is employed for pay,

${\beta}_{4}$ coefficient represents the change in log odds for respondent who is born in Canada,

${\beta}_{5}$ coefficient represents the change in log odds for respondent age between 30 to 44,

${\beta}_{6}$ coefficient represents the change in log odds for respondent age between 45 to 59,

${\beta}_{7}$ coefficient represents the change in log odds for respondent age between 60 to 74,

${\beta}_{8}$ coefficient represents the change in log odds for respondent age from 75 and above,

$x$ represents each factor respectively.

# 4 Results
## 4.1 Model Result
```{r,include=FALSE}
#Multilevel Logistic Model for different states
model <- glmer(vote_Conservative ~ sex + edu + employed + bornin_canada + as.factor(age) + (1|province), data=survey, family =binomial)
```

Multilevel logistic regression model estimates by interpreting regression coefficients:
$$\log \left(\frac{\hat{p}_{ij}}{1-\hat{p}_{ij}}\right)=0.37+0.57 x_{ij}^{sexMale}-0.66 x_{ij}^{educationed}+0.01 x_{ij}^{employed}-0.24 x_{ij}^{bornCA}$$
$$-0.05 x_{ij}^{age30-44}+0.24 x_{ij}^{age45-59}-0.02 x_{ij}^{age60-74}+0.14 x_{ij}^{age75above}$$
```{r,echo=FALSE}
#Statistical summary
kable(summary(model)[["coefficients"]], digits=4, 
      caption="Summary of Model Estimates")
```

## 4.2 Post-stratification Calculation Result
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#post-stratification calculation
#Making prediction using our model with census data, and we saving the results of those predictions by adding a new column to the census_data data set called 'estimate'
post$estimate <- 
  model %>% 
  predict(newdata = post, type="response")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
##together vote outcome
#together <- post %>%
#mutate(proportion_voting_for_conservative = estimate*n) %>%
#  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
#kable(together, digits=4, 
#      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted")

##summarize by geography
#geography <- post %>%
#mutate(proportion_voting_for_conservative = estimate*n) %>%
#  group_by(state) %>%
#  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
#kable(geography, digits=4, 
#      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted by Province")

```
Errors exists in the code for getting vote outcome. To be fixed.
